TG2 Publications

### 2025-08-18 papers.qmd
Added short sentence explaining why it is useful for each paper.

* Kipruto, E., & Sauerbrei, W. (2025). Evaluating Prediction Performance: A Simulation Study Comparing Penalized and Classical Variable Selection Methods in Low-Dimensional Data. Applied Sciences, 15(13), 7443. [https://doi.org/10.3390/app15137443](https://doi.org/10.3390/app15137443)

* Sauerbrei, W., Royston, P., & Kipruto, E. (2025). Multivariable Fractional Polynomial Models and Extensions. In International Encyclopedia of Statistical Science (pp. 1609-1616). Berlin, Heidelberg: Springer. [https://doi.org/10.1007/978-3-662-69359-9_397](https://doi.org/10.1007/978-3-662-69359-9_397)

* Kipruto, E., & Sauerbrei, W. (2024). Post‐Estimation Shrinkage in Full and Selected Linear Regression Models in Low‐Dimensional Data Revisited. Biometrical Journal, 66(7), e202300368. [https://doi.org/10.1002/bimj.202300368](https://doi.org/10.1002/bimj.202300368)

* Lu Z, Lou W. Bayesian approaches to variable selection: a comparative study from practical perspectives. Int J Biostat 2021; [https://doi.org/10.1515/ijb-2020-0130](https://doi.org/10.1515/ijb-2020-0130)

* Heinze G, Wallisch C, Dunkler D. Variable selection – a review and recommendations for the practicing statistician. Biometrical J. 2018; 60:431–49. 
* Binder H, Sauerbrei W, Royston P. Comparison between splines and fractional polynomials for multivariable model building with continuous covariates: a simulation study with continuous response. Statistics in Medicine 2013; 32(13): 2262– 2277.

* Sauerbrei W, Royston P, Binder H. Selection of important variables and determination of functional form for continuous predictors in multivariable model building. Statistics in Medicine 2007; 26: 5512– 5528.

* Abrahamowicz M, MacKenzie TA. Joint estimation of time‐dependent and non‐linear effects of continuous covariates on survival. Statistics in Medicine 2007; 26(2): 392– 408.

* Abrahamowicz M, Du Berger R, Grover SA. Flexible modeling of the effects of serum cholesterol on coronary heart disease mortality. American Journal of Epidemiology 1997; 145(8): 714– 729.

* Greenland S. Avoiding power loss associated with categorization and ordinal scores in dose‐response and trend analysis. Epidemiology (Cambridge, Mass.) 1995; 6(4): 450– 454.

* Royston P, Altman DG. Regression using fractional polynomials of continuous covariates: parsimonious parametric modelling. Applied Statistic 1994; 43(3): 429– 467.

>>>>

* Kipruto, E., & Sauerbrei, W. (2025). Evaluating Prediction Performance: A Simulation Study Comparing Penalized and Classical Variable Selection Methods in Low-Dimensional Data. Applied Sciences, 15(13), 7443. [https://doi.org/10.3390/app15137443](https://doi.org/10.3390/app15137443)
<br> *Highlights how the relative performance of penalized and classical selection methods depends on data conditions, guiding informed method choice*.

* Sauerbrei, W., Royston, P., & Kipruto, E. (2025). Multivariable Fractional Polynomial Models and Extensions. In International Encyclopedia of Statistical Science (pp. 1609-1616). Berlin, Heidelberg: Springer. [https://doi.org/10.1007/978-3-662-69359-9_397](https://doi.org/10.1007/978-3-662-69359-9_397)
<br> *Summarizes fractional polynomial modeling as a structured and practical approach to functional form selection.*

* Kipruto, E., & Sauerbrei, W. (2024). Post‐Estimation Shrinkage in Full and Selected Linear Regression Models in Low‐Dimensional Data Revisited. Biometrical Journal, 66(7), e202300368. [https://doi.org/10.1002/bimj.202300368](https://doi.org/10.1002/bimj.202300368)
<br> *Demonstrates how post-estimation shrinkage can reduce overfitting and bias, offering an alternative to penalized regression after variable selection*.

* Lu Z, Lou W. Bayesian approaches to variable selection: a comparative study from practical perspectives. Int J Biostat 2021; [https://doi.org/10.1515/ijb-2020-0130](https://doi.org/10.1515/ijb-2020-0130)
<br> *Provides practical comparisons of Bayesian variable selection approaches, supporting their application in real-world data analysis*.

* Heinze G, Wallisch C, Dunkler D. Variable selection – a review and recommendations for the practicing statistician. Biometrical J. 2018; 60:431–49. 
<br> *Offers pragmatic recommendations on variable selection methods, emphasizing model stability, validity, and transparency*.

* Binder H, Sauerbrei W, Royston P. Comparison between splines and fractional polynomials for multivariable model building with continuous covariates: a simulation study with continuous response. Statistics in Medicine 2013; 32(13): 2262– 2277.
<br> *Compares splines and fractional polynomials, clarifying their relative strengths for modeling continuous covariates*.

* Sauerbrei W, Royston P, Binder H. Selection of important variables and determination of functional form for continuous predictors in multivariable model building. Statistics in Medicine 2007; 26: 5512– 5528.
<br> *Integrates strategies for variable selection and functional form choice, presenting fractional polynomials as a coherent solution*.

* Abrahamowicz M, MacKenzie TA. Joint estimation of time‐dependent and non‐linear effects of continuous covariates on survival. Statistics in Medicine 2007; 26(2): 392– 408.
<br> *Illustrates how joint modeling of time-varying and nonlinear effects improves accuracy and avoids bias in survival analysis*.

* Abrahamowicz M, Du Berger R, Grover SA. Flexible modeling of the effects of serum cholesterol on coronary heart disease mortality. American Journal of Epidemiology 1997; 145(8): 714– 729.
<br> *Shows how flexible modeling of continuous exposures can reveal nonlinear risk relationships that parametric models may obscure*.

* Greenland S. Avoiding power loss associated with categorization and ordinal scores in dose‐response and trend analysis. Epidemiology (Cambridge, Mass.) 1995; 6(4): 450– 454.
<br> *Demonstrates the loss of power and information from categorizing continuous exposures, reinforcing the need for flexible functional form modeling*.

* Royston P, Altman DG. Regression using fractional polynomials of continuous covariates: parsimonious parametric modelling. Applied Statistic 1994; 43(3): 429– 467. 
<br> *Introduces fractional polynomials as a parsimonious and interpretable alternative to categorization or high-order polynomials*.
